---
title: "Case Study Projec: How Does a Bike-Share Navigate Speedy Success?t"
author: "Ayo"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Table of Contents
<ul>
<li><a href="#intro">Introduction</a></li>
<li><a href="#Ask">Aim and Objectives (ASK)</a></li>
<li><a href="#bt">Business Tasks</a></li>
<li><a href="#conclusions">Conclusions</a></li>
</ul>

<a id='intro'></a>
### Introduction
In 2016, Cyclistic launched a successful bike-share offering. Since then, the program has grown to a fleet of 5,824 bicycles that are geo tracked and locked into a network of 692 stations across Chicago. The bikes can be unlocked from one station and returned to any other station in the system anytime. Until now, Cyclistic’s marketing strategy relied on building general awareness and appealing to broad consumer segments. One approach that helped make these things possible was the flexibility of its pricing plans: single-ride passes, full-day passes, and annual memberships. Customers who purchase single-ride or full-day passes are referred to as casual riders. Customers who purchase annual memberships are Cyclistic members.Cyclistic’s finance analysts have concluded that annual members are much more profitable than casual riders. Although the pricing flexibility helps Cyclistic attract more customers, Moreno believes that maximizing the number of annual members will be key to future growth. Rather than creating a marketing campaign that targets all-new customers, Moreno believes there is a very good chance to convert casual riders into. She notes that casual riders are already aware of the Cyclistic program and have chosen Cyclistic for their mobility needs. The stakeholders has set a clear goal: Design marketing strategies aimed at converting casual riders into annual members. In order to do that, however, the marketing analyst team needs to better understand how annual members and casual riders differ, why casual riders would buy a membership, and how digital media could affect their marketing tactics. The stakeholders are interested in analyzing the Cyclistic historical bike trip data to identify trends.

<a id='Ask'></a> 
## Aim and Objectives (ASK)
Design marketing strategies aimed at converting casual riders into annual members

According to the ask phase of data analysis process three questions was generated which leads
to the three main objectives which are 
1. How do annual members and casual riders use Cyclistic bikes differently?
2. Why would casual riders buy Cyclistic annual memberships?
3. How can Cyclistic use digital media to influence casual riders to become members?

<a id='bt'></a>
### Business Tasks
As a junior data analyst working in the marketing analyst team at Cyclistic, a bike-share company in Chicago. The director of marketing believes the company’s future success depends on maximizing the number of annual memberships. Therefore, my team wants to understand how casual riders and annual members use Cyclistic bikes differently. From these insights,
my team will design a new marketing strategy to convert casual riders into annual members. But first, Cyclistic executives must approve your recommendations, so they must be backed up with compelling data insights and professional data
visualizations.

### My Personal Business Task
How do annual members and casual riders use cyclistic bike diferently?

### Stakeholders
>- Lily Moreno: The director of marketing and your manager. Moreno is responsible for the development of campaigns
and initiatives to promote the bike-share program. These may include email, social media, and other channels.
>- Cyclistic marketing analytics team: A team of data analysts who are responsible for collecting, analyzing, and reporting data that helps guide Cyclistic marketing strategy. You joined this team six months ago and have been busy learning about Cyclistic’s mission and business goals — as well as how you, as a junior data analyst, can help Cyclistic achieve them.
>- Cyclistic executive team: The notoriously detail-oriented executive team will decide whether to approve the recommended marketing program.

## PREPARE

### Dataset

For the purposes of this case study, the dataset appropriate and will enable me to answer the business questions is the cyclistic trip bike-share which is publicly available for download [here](https://divvy-tripdata.s3.amazonaws.com/index.html). This public data can  be used to explore how different customer types are using Cyclistic bikes. 

For data-privacy I won't be using riders's persnal identifiables information. This
means that i won’t be able to connect pass purchases to credit card numbers to determine if casual riders live in the Cyclistic service area or if they have purchased multiple single passes which will be one of my limitations.

Data is in comma seperated value (csv) format, downloaded from amazonaws website and stored in a directory in my local disk. To verify data integrity, proper data cleaning was done on the data set 

## Processing Data

### Importing the packages

```{r}
library(tidyverse)
library(janitor)
library(skimr)
library(lubridate)
```

### Loading the dat 
```{r}
df <- read.csv('E:/Alx-project3/202205-divvy-tripdata/202205-divvy-tripdata.csv')
df_copy <- df
```

### Assessing the data

```{r}
head(df, 5)
```
```{r}
str(df)
colnames(df) # getting the column names in the dataframe
glimpse(df)
skim_without_charts(df)
```

### Cleaning
```{r}
df_copy$start_date <- as.Date(df_copy$started_at) 
df_copy$start_month <- format(as.Date(df_copy$start_date), "%m")
df_copy$start_day <- format(as.Date(df_copy$start_date), "%d")
df_copy$start_year <- format(as.Date(df_copy$start_date), "%Y")
df_copy$start_week <- format(as.Date(df_copy$start_date), "%A")
df_copy$end_date <- as.Date(df_copy$ended_at)
```

```{r}
skim_without_charts(df)
```

```{r}
# dropping some unwanted columns
colnames(df_copy)
df_copy <- df_copy %>% select(-c(start_lat, start_lng, end_lat, end_lng))
```

```{r}
# Creating new features ride_duration
df_copy$ride_duration <- difftime(df_copy$ended_at, df_copy$started_at)
str(df_copy)
skim_without_charts(df_copy)
```
Seems we have negative ride duration, so lets remove that

```{r}
df_clean <- filter(df_copy, df_copy$ride_duration >= 0)
```

Rename the start day of the week to be more descriptive
```{r}
df_copy <- rename(df_copy, start_day_of_week=start_week)
summary(df_clean)
skim_without_charts(df_clean)
table(df_clean$member_casual)

```
lets drop some irrelevant datas
```{r}
df_clean <- filter(df_copy, df_copy$ride_duration >= 0)
```


Convert ride duration to seconds
```{r}
df_clean$ride_duration <- as.numeric(df_clean$ride_duration)
is.numeric(df_clean$ride_duration)
```
 
 
### Analysis

```{r}
summary(df_clean$ride_duration)
```
Here we have the summary statistics of the ride duration regardless of member type
The total average ride duration is 1266secs
Minimun ride duration is 0sec
Maximun ride duration is 2175468secs


Lets gets the average duration for the members type 
```{r}
aggregate(df_clean$ride_duration ~ df_clean$member_casual, FUN = mean)
```
Casual riders seems to have significant high ride duration member riders than the member riders

```{r}
head(df_clean)
```

```{r}
df_clean$start_day_of_week <- ordered(df_clean$start_day_of_week, levels=c('Sunday', 'Monday', 
                                                                           'Tuesday',
                                                                           'Wednessday', 'Thursday', 
                                                                           'Friday',
                                                                           'Saturday'))
```


```{r}
aggregate(df_clean$ride_duration ~ df_clean$member_casual + df_clean$start_day_of_week, FUN = mean)
```

